### 2021-05-07 Paper/Methodology.tex
latex now basically to the end!!
the final push :)

As discussed in the literature review, there are numerous models for assessing asset pricing and abnormal returns. In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model". This analysis method is built on the Fama French three-factor model and relies on the two-pass sort. Like the section \ref{Data} on data, this section is broken up into multiple parts. First, I will explain the methodology for assessing the sensitivity and risk premium for the ESG factor created from the MSCI ESG index minus the petroleum index. Next, I will discuss the methodology for MSCI ESG rating ESG factors. I will conclude with a brief description of the newey west errors and the setup of assessing statistical significance.

For the first part of this study, I sort each security into decile groups based on beta and size. The size and beta deciles ratings are provided by CRSP and are merged on the individual firm level for each month of data. Next, I conducted the first pass for the two-pass sort by running the regression of expected returns on market risk premia the four Fama French factors and the EMP factor as described in equation \eqref{six_factor_basic_reg}. This regression was run twice: once on the beta-ranked portfolios and again on the size-ranked portfolios.

\begin{equation}
\begin{split}
    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ h_iHML_t+  r_iRMW_t+c_iCMA_t+e_iEMP_t +\epsilon_{i,t}
    \end{split}
    \label{six_factor_basic_reg}
\end{equation}

\begin{center}
Where:\\
    $R_{i,t}^e = R_{it}-R_{rf,t}$
\end{center}

Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort. The second pass of the two-pass sort is described in the equation  \eqref{second_pass}. The second pass describes the risk premia for each of the six factors in the mode. This second pass was performed for both beta and size-based portfolios.

\begin{equation}
    E[R^e]= b\lambda_1+s\lambda_2 +h\lambda_3 +r\lambda_4 +c \lambda_4 + e\lambda_5
    \label{second_pass}
\end{equation}\\

The second set of regressions is focused on establishing a link between MSCI ESG ratings and ESG risk factor sensitivity. Similar to the first set of regressions, I use a two-pass sort methodology using both size and beta ranked portfolios. To create the ESG factor for this set of analyses, I used the all industry ESG factor described in Section \ref{Data}. The process of implementing the regression follows the same basic form as the one for the MSCI ESG index and petroleum index ESG factor by first running a pass on the portfolio by size and following that with a cross-sectional analysis to quantify the risk premium.

Finally, in determining the statistical significance of each of the coefficients, it is important to discuss the newly west error. The newly west error is a replacement for standard error in a traditional OLS regression. the advantage of the newly west is an ability to account for serial correlation within the model

>>>>

As discussed in the literature review, there are numerous models for assessing asset pricing and excess returns. In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model". This analysis method is built on the Fama French three-factor model and relies on the two-pass sort. First, I will explain the methodology for assessing the sensitivity and risk premium for the ESG factor created from the MSCI ESG index minus the petroleum index. Next, I will discuss the methodology for MSCI ESG rating ESG factors. I will conclude with a brief description of the newey west errors and the setup of assessing statistical significance.

For the first part of this study, I sort each security into decile groups based on beta and size. The size and beta deciles ratings are provided by CRSP and are merged on the individual firm level for each month of data. Next, I conducted the first pass for the two-pass sort by running the regression of excess returns on market risk premia, the four Fama French factors, and the EMP factor as described in equation \eqref{six_factor_basic_reg}. This regression was run twice: once on the beta-ranked portfolios and again on the size-ranked portfolios. The first pass provides insight into the sensitivity of the factor on the declie index. 
\begin{equation}
\begin{split}
    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ h_iHML_t+  r_iRMW_t+c_iCMA_t+e_iEMP_t +\epsilon_{i,t}
    \end{split}
    \label{six_factor_basic_reg}
\end{equation}
\begin{center}
Where:\\
    $R_{i,t}^e = R_{it}-R_{rf,t}$
\end{center}
Next, I used the output data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort as described in the equation  \eqref{second_pass}. The second pass represents the risk premia for each of the six factors in the model. This second pass was performed for both beta and size-based portfolios.
\begin{equation}
    E[R^e]= b\lambda_1+s\lambda_2 +h\lambda_3 +r\lambda_4 +c \lambda_4 + e\lambda_5
    \label{second_pass}
\end{equation}\\
The second set of regressions is focused on establishing a link between MSCI ESG ratings and excess returns. Similar to the first set of regressions, I use a two-pass sort methodology using both size and beta ranked portfolios. To create the ESG factor for this set of analyses, I used the ESG factor described in Section \ref{Data}. The process of implementing the regression follows the same basic form as the one for the MSCI ESG index and petroleum index ESG factor by first running a pass on the portfolio by size(beta) decile and following that with a cross-sectional analysis to quantify the risk premium.

Finally, in determining the statistical significance of each of the coefficients, it is important to discuss the newly west error. The newly west error is a replacement for standard error in a traditional OLS regression. The advantage of the newly west is an ability to account for serial correlation within the model. To view any of the supporting code for the analysis, view the link in the citation \cite{Matt_code}.

### 2021-05-07 Paper/Results.tex
the final push :)

Table 2 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}. The row axis is the distinct size deciles, where 1 is the smallest 10\% of companies by market cap companies, and 10 is the decile with the largest 10\% of companies by market cap. The column axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor. $HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.  $SMB_{t}$ is the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks. $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability. $CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms. $EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index. Finally, the intercept is the excess returns for the given decile. The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor. In addition, the EMP factor also exhibits a consistent negative relationship between the EGS factor and returns. This relationship indicates that investors are returning a premium by investing in ESG companies over the contra ESG counterparts and that this relationship is largest for small companies. The results from this statistical test are significant at the 5\% level for the EMP factor for all of the portfolios except the smallest and largest size portfolios. However, several of the Fama French factors are also not significant for the largest and smallest portfolios, which may indicate that these tail portfolios might be behaving differently due to their composition rather than a reflection of the factor.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/size_output_inital}\\
    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size. the row axis is portfolio and the column axis is the  market beta,four Fama French factors and the EMP factor\\
    \paperspacingwide
\end{center}

Table 3 demonstrates the risk premia for each of the factors using the second step in the two-pass sort for Size decile portfolios. The market factor, the SMB factor, the CMA factor, and the EMP factor are all positive and statistically significant at the 1\% level. The other two factors have a negative sign and are not statistically significant at the 10\% level.  EMP factor demonstrates that a 1 unit change in beta corresponds with an 8.7\% increase in returns. Although this is by far the smallest coefficient of the four statistically significant returns, this term is still economically meaningful and represents a risk premia for investing in high ESG firms. Overall the model appears to be a fairly good fit with a $R^2$ value of 87.5\%, which is lower than some previous literature using the petroleum index but still indicates a well fit model \hl{cite this paper}

\begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2}\\
    \textbf{Table 3:} shows the results for the Second pass of the twp pass sort. On the left labeled as (1) are the regression results for the size decile portfolios. On the right labeled as (2) are the regression results for the beta decile portfolios\\
    \paperspacingwide
\end{center}

Table 4 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the row axis is the beta decile. In this table,  the beta decile 1 comprises the companies with the lowest market beta's. Similarly, beta decile 10 is the portfolio of companies with the largest market beta values.  Unlike the size decile portfolios, the beta decile portfolios do not have consistently negative relations with respect to decile. Rather the relationship appears to have a dome shape where portfolio 6 has the smallest positive sensitivity to the EMP factor, and larger or smaller beta portfolios appear to have a larger sensitivity to the  EMP factor. \hl{why might this be}. Similar to the size decile analysis, the EMP coefficients are all very significant, with every portfolio having significance at the 5\% level.

When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB, CMA, and EMP exhibit positive risk premia. The other two factors remain statistically insignificant, and the MHL factor changed signs and now has a positive relationship. The EMP coefficient indicates that a one-unit change in beta would result in a 7.36\% change in returns for a high ESG company. This result is slightly smaller than the result from the size portfolio but still demonstrates a statistical relationship in the same order of magnitude. The model is also a good fit with an $R^2$ value of 86.0\%. With the baseline analysis proving that there is a relationship between ESG and excess marker returns, I will provide an investigation of the robustness in section \ref{Robustness} by first comparing the results of the two-pass sort with the EGS factor to the same analysis without the factor. Next, I will investigate time-related relationships between ESG and excess returns by implementing a dummy variable for the year and repeating the two-pass sort methodology.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/beta_output_inital}\\
    \textbf{Table 4:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
    \paperspacingwide
\end{center}

Moving on to the second stage of the analysis, I constructed an ESG factor out of the MSCI ESG ratings to capture a more holistic view of ESG risk. I created this factor by taking the top 10\% of MSCI EGS rated companies and subtracted them from the bottom 10\% of firms rated by the MSCI. I repeated the methodology above for the regression analysis. I  regressed two sets of two-pass regression analyses, one on beta-sized portfolios and another on size-based portfolios. In the first pass results for the MSCI factor, the beta portfolio demonstrated a decrease in sensitivity as the betas grew; however, the size-based portfolios did not experience any real pattern governing the decile and the sensitivity. In the case of the size-based portfolios, some of the sensitivity figures were positive while others were negative. Table 5 represents a second stage two-pass sort regression output. Similar to the regression in the first stage of the analysis, the market return, CMA, SMB remain statistically significant; however, only the size-based portfolios demonstrate a statistically significant coefficient for the ESG factor.  The EMP ESG factor coefficient can be interpreted as a negative relationship between ESG and excess return wherein a one beta unit increase in ESG results in a decrease of 2\% in excess returns. One limitation of the MSCI data is it only represents a snapshot in ESG rating data; that is, the MSCI data is only the most recent ESG performance and does not account for changes in ESG performance. As such, a company that used to be a  poor ESG and has improved to a strong ESG company will only be categorized as a strong ESG company which will not accurately address the ESG performance.  As such, in section \ref{Robustness} I will address this by shrinking the time frame to see if the coefficients increase in their significance. Further,  to address additional concerns of robustness, I will run another two-pass sort to isolate the SIC industry-specific ESG factor and determine if there are ESG related excess returns that are greater in some industries than others.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/All_Industry_FF_pass_2}\\
    \textbf{Table 5:} SIC top 10\% minus bottom 10\% \hl{NEED TO EDIT THESE RESULTS}
    \paperspacingwide
\end{center}

>>>>

Table 2 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}. The row axis is the distinct size deciles, where 1 is the smallest 10\% of companies by market cap companies, and 10 is the decile with the largest 10\% of companies by market cap. The column axis includes estimated coefficients for the five Fama French factors and the EMP factor. $HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.  $SMB_{t}$ is the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks. $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability. $CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms. $EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index. The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor.  This relationship indicates that investors are returning a premium by investing in ESG companies over the contra ESG counterparts and that this relationship is strongest for small companies.  The size-based sensitivity that is greatest for small firms was also noted in Dorfleitner et al's work\cite{Dorfleitner2020ESGCA}. The results from this statistical test are significant at the 5\% level for the EMP factor for all of the portfolios except the smallest and largest size portfolios. However, several of the Fama French factors are also not significant for the largest and smallest portfolios. This finding indicate that these tail portfolios  behave differently and may require winsorizing rather than a reflection of the factor's significance.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/size_output_inital}\\
    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size deciles. The row axis is portfolio and the column axis is the  market beta, four Fama French factors, and the EMP factor.\\
    \paperspacingwide
\end{center}
Table 3 demonstrates the risk premia for each of the factors using the second step in the two-pass sort for size decile portfolios. The market factor, the SMB factor, the CMA factor, and the EMP factor are all positive and statistically significant at the 1\% level. The other two factors have a negative sign and are not statistically significant at the 10\% level.  EMP factor demonstrates that a 1 unit change in beta corresponds with an 8.7\% increase in excess returns. Although this is by far the smallest coefficient of the four statistically significant factors, this term is still economically meaningful and represents a risk premium can be achieved by investing in high ESG firms. Overall the model appears to fit the data reasonably well with a $R^2$ value of 87.5\%, which is lower than some previous literature using the petroleum index but still indicates a well-fit model \cite{Gregory2020TheGP}.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/FF_pass_2}\\
    \textbf{Table 3:} shows the results for the second pass of the two pass sort with the EMP factor. On the left labeled as (1) are the regression results for the size decile portfolios. On the right labeled as (2) are the regression results for the beta decile portfolios.\\
    \paperspacingwide
\end{center}

Table 4 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the row axis is the beta decile. In this table,  the beta decile 1 comprises the companies with the lowest market beta's. Similarly, beta decile 10 is the portfolio of companies with the largest market beta values.  Unlike the size decile portfolios, the beta decile portfolios do not have consistently negative relations with respect to decile. Rather the relationship appears to have a dome shape where portfolio 6 has the smallest positive sensitivity to the EMP factor, and larger or smaller beta portfolios appear to have a larger sensitivity to the  EMP factor. Similar to the size decile analysis, the EMP coefficients are all highly significant, with every portfolio having significance at the 5\% level.

When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB, CMA, and EMP exhibit positive risk premia and the other two factors remain statistically insignificant. The EMP coefficient indicates that a one-unit change in beta would result in a 7.36\% change in excess returns for a high ESG company. This result is slightly smaller than the result from the size portfolio but still demonstrates a statistical relationship in the same order of magnitude and suggests the EGS factor can provide excess returns. The beta decile model is a strong fit with an $R^2$ value of 86.0\%.

The two-pass sort indicates that there is a measurable relationship between the EMP factor and excess marker returns. In section \ref{Robustness} I will provide an analysis of the robustness by first comparing the results of the two-pass sort with the EGS factor to the same analysis without the factor. Next, I will investigate the time-related relationships between ESG and excess returns by implementing a dummy variable for the year and repeating the two-pass sort methodology.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/beta_output_inital}\\
    \textbf{Table 4:} shows the first pass in a two pass sort. The portfolios are decomposed by beta deciles and the table demonstrates the sensitivity of the four factors, market returns and and EMP factor to the decile.\\
    \paperspacingwide
\end{center}

Moving on to the second stage of the analysis, I constructed an ESG factor out of the MSCI ESG ratings to capture a more holistic view of ESG risk. I created this factor by taking the top 10\% of MSCI EGS rated companies and subtracted them from the bottom 10\% of firms rated by the MSCI. I repeated the methodology above for the regression analysis. I  regressed two sets of two-pass sort  regression analyses, one on beta-sized portfolios and another on size-based portfolios. In the first pass results for the MSCI ESG factor, the beta portfolio demonstrated a decrease in sensitivity as the beta deciles increased. However, the size-based portfolios did not experience any pattern, with some of the coefficients were positive while others were negative.

Table 5 represents the second stage in two-pass sort regression output. Similar to the regression with the EMP factor, the market return, CMA, SMB remain statistically significant. However, neither the size based size-based or beta-based portfolios demonstrate a statistically significant coefficient for the ESG factor. This lack of statistical significance indicates that the MSIC ESG factor is priced into one of the other statistically significant factors and should not produce any excess returns. Further solidifying that the factor is not economically significant is the high $R^2$  value for both of the portfolios, which indicates the strong fit of the model.

One potential reason that there is no statistical significance for the factor is the limitation of the MSCI data. MSCI only provides current  MSCI ESG rating data. As such, a company that used to be a  poor ESG firm that has improved to a strong ESG company will only be categorized as a strong ESG company which will not accurately address the ESG performance.

\begin{center}
    \paperspacingnarrow

    \input{Code/data/All_Industry_FF_pass_2}\\
    \textbf{Table 5:} shows the results for the second pass of the two-pass sort with the ESG factor generated from the top 10\% of MSCI ESG scores minus bottom 10\% MSCI ESG scores. On the left labeled as (1) are the regression results for the size decile portfolios. On the right, labeled as (2) are the regression results for the beta decile portfolios.

    \paperspacingwide
\end{center}

In section \ref{Robustness} I will address the data concerns by reducing the time frame and observing if the coefficients increase in significance. Further, to address additional concerns of robustness, I will run another two-pass sort on individual industries to see if there is uniformity to the risk prima response to the MSCI ESG factor.

### 2021-05-07 Paper/conclusion.tex
the final push :)

Looking beyond this paper's scope and future research opportunities, there are a few natural extensions for assessing the role rating agencies and indexes play in the EGS landscape. One extension is to further isolate ESG premia by conducting an event study methodology on several well-known ESG indexes, including the MSCI ESG leaders index referenced in this paper. Using addition and deletion events from this index can provide a tangible event to test the market reaction to changes in ESG news and assess if there are abnormal returns. This methodology for evaluating abnormal returns can also be applied to ESG rating; however, there might be less of an effect if the market was anticipating a change in the rating. As such, the indexing market event might provide a more precise image regarding market tracking behavior around ESG.

Another possible extension for ESG research is to take a further look at the rating agencies. Over the last few years, more companies have entered into the ESG rating space, MSCI and Sustainalytics have held a dominant position in the market, but S\&P, Moody's, and other firms are looking to expand their research efforts around ESG research. An interesting extension would be to analyze the risk premia for different firms and see if there an ESG factor where high ESG minus low ESG has a larger risk premium for one rating agency over another rating agency.

>>>>

The main conclusions from this paper come from defining and implementing an ESG factor.  In the first set of analyses where the ESG term is defined as the difference between the MSCI ESG index and the petroleum index, I observed a statistically significant result for the ESG term, which matched previous work conducted on ESG factor modeling. However, in the second set of analyses, when developing an ESG factor from the MSCI ratings, I did not find a statistically significant factor on the aggregate level. Examining the MSCI results at the one-digit SIC level, I was only able to find the firms starting with SIC code  1  to demonstrate a positive and statistically significant factor. These findings reaffirm that the petroleum industry may have negative expected returns but does not build a case for ESG providing excess returns for other industries. This lack of uniformity across both definitions fosters a number of important questions to address in the implementation of the Fama french six-factor model. This uncertainty in the findings of an ESG factor parody the literature review and make it extremely difficult to recommend a pro ESG policy proposal.

Policy for EGS investing over the last year has been a contentious topic of debate. In 2020, the Department of Labor ruled that private pension plan providers ``cannot invest in ESG vehicles that sacrificed investment returns or take on additional risk" \cite{beals_trump_nodate}. Investment and retirement planning are essential to the wellbeing of everyone in their retirement, and 401(k) plans are a key part of securing the ability to continue a comfortable life into retirement.  Prioritizing returns over a longer period,  even by a few basis points can have a sizable impact when considering the effect of compounding over a time frame of 40 years. When comparing the current policy position to the results from this study, it would seem like a clear solution to uphold the policy laid out by the Department of Labor. However, the results suggest that investors would be better off if pension providers actively avoided traditional fossil fuels as the core of energy production and utilization transitions to renewable solutions.

Further, it is important to note that the results from this study are retrospective and do not account for future development. Under the Biden administration, there has been a shift in political tone regarding environmental and social concerns that would support corporate ESG initiatives \cite{newburger_heres_2021}. As such, many industries might benefit from these new political tailwinds and create industry ESG divergence similarly to those observed in the petroleum industry.

To account for this uncertainty and opportunity,  I would recommend rewording the mandate from the Bureau of Labor to be more lenient and offer investment solutions that incorporate ESG for those customers who have an active interest in financing global environmental and social reform. However,  these investment products should also be provided with additional warnings for the inventor and strict regulations regarding performance claims. Return performance should demonstrate the excess return or loss associated with the product's historical performance compared to a not ESG counterpart fund.

Looking beyond this paper's scope to future research opportunities, there are a few natural extensions for assessing the role rating agencies and indexes play in the EGS landscape. One extension is to further isolate ESG premia by conducting an event study methodology on several well-known ESG indexes, including the MSCI ESG leaders index referenced in this paper. Using addition and deletion events from this index can provide a tangible event to test the market reaction to changes in ESG news and assess if there are abnormal returns. This methodology for evaluating abnormal returns can also be applied to ESG rating changes; however, there might be less of an effect if the market was anticipating a change in the rating similar to corporate credit ratings. As such, the indexing market event might provide a more precise image regarding market tracking behavior around ESG.

Another possible extension for ESG research is to take a further look at the rating agencies. Over the last few years, more companies have entered into the ESG rating space, MSCI and Sustainalytics have held a dominant position in the market, but S\&P, Moody's, and other firms are looking to expand their research efforts around ESG research. An interesting extension would be to analyze the risk premia for different firms and see if there is an ESG factor where high ESG minus low ESG has a larger risk premium for one rating agency over another rating agency.

### 2021-05-07 Paper/data.tex
Updates from Overleaf
the final push :)

The data used for this paper was sourced from four datasets. The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices. Data for the daily returns of MSCI World ESG Focus Index and S\&P petroleum index was sourced from CAPIQ Compustat. Data on firm-level ESG scores (by ticker) was sourced from MSCI and reflected the most recent company ratings as MSCI does not have the historical ratings relatively available. Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R. French - Data Library.

The data and methodology for the paper are broken into two sections; first, a preliminary study is conducted using a broad base ESG factor to confirm a statistically risk sensitivity relationship between ESG and returns. Next, an analysis on a SIC level is conducted in an attempt to understand how industry-specific ESG factors affect the risk-return relationship of stock performance.

In the first stage of the analysis, an ESG factor is created by taking two market indexes and subtracting their daily returns in an attempt to isolate ESG based risk. The formula for this factor is represented in Equation \eqref{making_EMP}, and for the remainder of the section, one analysis will be referred to as EMP, which is sort for ESG minus petroleum. This method of creating a factor is very similar to Fama and French in their other four factors in the five-factor model. For this analysis, I will be using the MSCI World ESG Focus Index as the high ESG portfolio and the  S\&P petroleum index as the low ESG portfolio. The methodology for creating an ESG factor is similar to the work from \hl{put the author's name in here}. In \hl{put the author's name in here}'s work, they were able to find a statistical risk relationship between a high ESG portfolio and a petroleum index\cite{Frynas2005TheFD}. The petroleum index provides a good contra ESG portfolio as the petroleum industry is in direct opposition to environmental goals that are E components of ESG. A similar analysis could also be run with other market indexes. The S\&P Mining index is another popular index that also performs very poorly for environmental and social concerns\cite{Kapelus2002MiningCS}. A final index that also could represent a contra ESG portfolio is the  \$VICE ETF. This financial product tracks the market for  alcohol, drug, gaming, and betting industry which have all been shown as clear contra ESG industries \hl{inset citation}
\begin{equation}
    \label{making_EMP}
    EMP = R_{\text{World ESG Focus Index}}- R_{\text{S\&P petroleum index}}
\end{equation}

To conduct this analysis, daily data was collected between June 2011 and March 2021.  June 2011 marks the initiation of the MSCI ESG index; as such, this is the beginning of the sample period for this paper. Daily individual stock return information was collected from CRSP and merged with the daily ESG factor returns. The return data was collapsed on month to create a monthly return figure by summing all of the daily returns for each given month in the dataset. Next, the Fama French monthly factors were merged into the dataset for each individual firm for each month. After aggregating, the factor and individual firm return, there were 830,891 observations of monthly return data for 12,194 unique stocks. The merged and cleaned dataset had an average of 68 months of data for each individual security. Table 1 provides brief summary statistics for the data in the first stage of the analysis.

In the second stage of the analysis, I look at how SIC code-specific ESG factors affect the risk-return relationship. In this study, I used the MSCI ESG rating by ticker to created SIC-based ESG factors. MSCI rate companies on their ESG score AAA-CCC where AAA is the highest ESG rating and CCC is the lowest ESG rating. In the sample of MSCI ratings, 16 companies were AAA, 88 companies were AA, 118 companies were A, 170 companies were BBB, 108 companies were BB, 40 companies were B, and 5 companies earned a CCC rating.  Each of these ratings was assigned a mating numeric score (1-7) and were merged using the CRSP name file to their corresponding kypermno value by matching ticker information. Subsequently, the ESG scores were merged into the monthly return dataset described at the end of the first analysis. A final process was conducted to make a unique ESG factor for each of the SIC codes. The unique ESG factor was created at the first digit SIC level and generated by splitting the SIC level ESG data into deciles and taking the top 10\% and as the high ESG index bottom 10\% as the low ESG and subtracting high ESG minus low ESG as explained in equation \eqref{making_EMP_section}. This ESG factor was created for SIC codes beginning with 1-4 and 6-8. SIC code. SIC code 5 only had BBB ratings, and MSCI did not have significant data on SIC code 0 or 9. In addition to creating industry specific ESG factors I also created a all industry ESG factor using the top 10\% and bottom 10\% for all MSCI ESG scores to create an ESG factor. \hl{explain how EMP is the same down here for consistency of reuslts}

\begin{equation}
    \label{making_EMP_section}
    EMP_i = R_{\text{high ESG,i}}- R_{\text{low ESG,i}}
\end{equation}

\begin{center}
    Where i is the first digit SIC code    
\end{center}

>>>>

The data used for this paper was sourced from four datasets. The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices. Data for the daily returns of MSCI World ESG Focus Index and S\&P petroleum index was sourced from CAPIQ Compustat. Data on firm-level ESG scores (by ticker) was sourced from MSCI and reflected the most recent company ratings. Finally, monthly Fama French factors for SMB, CMA, HML, and RMW were sourced from the \textit{Kenneth R. French Data Library}.

The data and methodology for the paper are broken into two sections; first, a preliminary study is conducted using a broad base ESG factor constructed from two market indexes to confirm a statistical relationship between ESG and excess returns. Next, another ESG factor is constructed using the MSCI ESG dataset  to isolate additional ESG risk.

In the first stage of the analysis, an ESG factor is created by taking two market indexes and subtracting their daily returns in an attempt to isolate ESG based risk. The formula for this factor is represented in Equation \eqref{making_EMP}, and for the remainder of this analysis, this ESG factor will be referred to as the EMP factor (ESG minus Petroleum). This method of creating a factor is based on the fama french model used by the other four factors in the five-factor model. For this analysis, I will be using the MSCI World ESG Focus Index as the high ESG portfolio and the  S\&P petroleum index as the low ESG portfolio. The choice of the petroleum index was based on Gregory et al's work where the authors demonstrated the sensitivity of the index to contra ESG factors \cite{Frynas2005TheFD}. The petroleum index provides a good contra ESG portfolio as the petroleum industry is in direct opposition to environmental goals that are `E' components of ESG. A similar analysis could also be run with other market indexes. The S\&P Mining index is another popular index that also performs very poorly for environmental and social concerns\cite{Kapelus2002MiningCS}. A third index that also could represent a contra ESG portfolio is the  \$VICE ETF. This financial product tracks the market for alcohol, drug, gaming, and betting industry which have all been shown as clear contra ESG industries \cite{Grougiou2016CorporateSR}.

\begin{equation}
    \label{making_EMP}
    EMP = R_{\text{World ESG Focus Index}}- R_{\text{S\&P petroleum index}}
\end{equation}

To conduct this analysis, daily data was collected between June 2011 and March 2021.  June 2011 marks the initiation of the MSCI ESG index; as such, this is the beginning of the sample period for this paper. Daily individual stock return information from CRSP was merged with the daily ESG factor returns. The resulting merged data was collapsed on months to create a monthly return figure by summing all of the daily returns for each given month in the dataset. Next, the Fama French monthly factors were merged onto the dataset for each individual firm for each month. After aggregating, the factor and individual firm return, there were 830,889 observations of monthly return data for 12,194 unique stocks. The merged and cleaned dataset had an average of 68 months of data for each individual security. Table 1 provides brief summary statistics for the data in the first stage of the analysis.

\begin{center}
    \paperspacingnarrow
    \input{Code/data/desc1}\\
    \textbf{Table 1:} shows summary statistics for the data set
    \paperspacingwide
\end{center}

In the second stage of the analysis, the definition of the ESG factor is changed to isolate additional ESG risk. In the first part of the analysis, the contra ESG portfolio's main concern centered on environmental risk. To increase exposure to social and governance while maintaining exposure to environmental risk, a portfolio is constructed from MSCI ESG ratings. MSCI provides public ratings for 546 firms and issues a rating from AAA-CCC where AAA is the highest ESG rating, and CCC is the lowest ESG rating. In the sample of MSCI ratings, 16 companies were AAA, 88 companies were AA, 118 companies were A, 170 companies were BBB, 108 companies were BB, 40 companies were B, and 5 companies earned a CCC rating.  Each of these ratings was assigned a matching numeric score (1-7) and was merged using the CRSP name file to their corresponding kypermno value by matching ticker information. The data was merged onto the monthly return dataset described at the end of the first analysis. To create an ESG factor with the MSCI rating, numeric ratings were split into deciles, and a high minus low  ESG factor was generated by taking the top 10\% of ESG firms and subtracting that from the bottom 10\%  of ESG firms as described in equation \eqref{making_EMP_section}. For the remainder of this analysis, the MSCI ESG factor will be referred to as the ESG factor or the MSCI ESG factor.

\begin{equation}
    \label{making_EMP_section}
    EMP = R_{\text{top 10\% ESG Firms}}- R_{\text{Bottom 10\% ESG Firms}}
\end{equation}
